Brownian motion of main-belt asteroids on human timescales

### 2025-02-16 main.tex
dN/dH done

\subsubsection{Encounter rates and geometries}
The pairwise interaction rate $dn/dA\,dv$ and the impulse direction
distributions $\langle \hat b^2_\gamma \rangle$ are estimated from a
numerical integration of all of the MBA orbits available
from the Minor Planet Center (MPC) as of ???? 2025.   We retain as
potential deflectors the 1.26~million bodies with $1.8<a<4.2$~AU,
observations at multiple oppositions, and uncertainty values $U\le5.$

From this full MBA list, we select a subset of 5000 objects as
``tracers'' randomly selected from the MBAs in each of three regimes:
``Inner'' MBAs with semi-major axes between the 4:1 and 3:1
mean-motion resonances of Jupiter; ``Middle'' MBAs between the 3:1 and
5:2 resonances; and ``Outer'' MBAs between the 5:2 and 2:1
resonances.  This is enough to be representative of the statistics of
each MBA class, but few enough that the calculation can be done easily
on a laptop computer.

@@

The leapfrog integrator advances all 1.26 million state vectors by 2
days, and the process is repeated.  We continue at 2-day intervals
until we have recorded 10 years' worth of encounters---17.7 million
total impulses.  These events are split according to their combination
of tracer class (Inner, Middle, or Outer belt) and deflector class
(Inner, Middle, Outer, Other).

For each type of impulse (\eg Inner-Inner), we combine the list of
events with the counts of candidate tracers and deflectors to
calculate $dn/dA\,dv,$ the event rates per eligible pair; and the mean
geometry of the encounter, $\langle \hat b^2_\gamma \rangle,$ as a
function of $v$.  Figure~\ref{fig:rtz} plots these functions of $v$
for the Inner-Inner, Middle-Middle, and Outer-Outer scattering
events.  Here it is apparent that the $\hat b_\phi$ component is
larger than $\hat b_z$ and $\hat b_r,$ but there is significant
variation with MBA class and with $v$.

@@

Asteroid counts $dN/dH$

>>>>

\subsubsection{Encounter rates and geometries}
The pairwise interaction rate $dn/dA\,dv$ and the impulse direction
distributions $\langle \hat b^2_\gamma \rangle$ are estimated from a
numerical integration of all of the MBA orbits available
from the Minor Planet Center (MPC) as of ???? 2025.   We retain as
potential deflectors the 1.26~million bodies with $1.8<a<4.2$~AU,
observations at multiple oppositions, and uncertainty values $U\le5.$
We designate each source as being an ``Inner'' MBA, with semi-major axes between the 4:1 and 3:1
mean-motion resonances of Jupiter; ``Middle'' MBA between the 3:1 and
5:2 resonances; ``Outer'' MBAs between the 5:2 and 2:1
resonances; and the small remainder as ``Other.''  Our working
assumption will be that the MPC objects are an unbiased sample of the
orbital-element distribution within each of the Inner, Middle, and
Outer belt regions.

From this full MBA list, we select a random subset of 5000 objects
from each of the Inner, Middle, and Outer belts to serve as a sample
of ``tracers.''  We record all of the passages of any tracer MBA
within 0.03~AU of any other ``deflector'' MBA (drawn from the full
catalog of 1.26~million) over a 10-year period.  The 5000 tracers per
belt region are enough to be representative of the statistics of
the region, but few enough that the calculation can be done easily
on a laptop computer.
We will report $\langle nI_\gamma^2\rangle$
separately for tracers of each region, allowing for the fact that a
tracer of a given region can receive impulses from deflectors resident
in all regions.

@@

The leapfrog integrator advances all 1.26 million state vectors by 2
days, and the process is repeated.  We continue at 2-day intervals
until we have recorded 10 years' worth of encounters---17.7 million
total impulses.

For each pair of tracer-deflector regions (\eg Inner-Inner), we combine the list of
events with the counts of candidate tracers and deflectors to
calculate $dn/dA\,dv,$ the event rates per eligible pair; and the mean
geometry of the encounter, $\langle \hat b^2_\gamma \rangle,$ as a
function of $v$.  Figure~\ref{fig:rtz} plots these functions of $v$
for the Inner-Inner, Middle-Middle, and Outer-Outer scattering
events.  Here it is apparent that the $\hat b_\phi$ component is
larger than $\hat b_z$ and $\hat b_r,$ but there is significant
variation with MBA class and with $v$.

@@

\begin{figure}
  \includegraphics{mbacounts.pdf}
  \caption{Differential counts of MBAs cataloged by the MPC are
    plotted vs absolute magnitude $H$  
    for the Inner, Middle, Outer, and ``Other'' regions of
    the main belt, and for their total.  We assume that they are
    complete for $H\le17,$ and extrapolate to fainter sources using
    the functional form from \eqq{eq:lsstmba} shown as the dotted
    curve.}
  \label{fig:counts}
\end{figure}

Figure~\ref{fig:counts} plots our assumptions for $dN/dH.$  The MPC
catalog is likely close to complete in all three regions for $H\le17,$
so we will use the MPC counts directly in this regime.  Note that the
three regions have similar shapes of $dN/dH.$  For $H>17,$ we adopt
the functional form for the cumulative MBA counts given by
initial LSST predictions \citep{LSST_science_book}:
\begin{equation}
  N(<H) \propto \frac{10^{0.43(H-15.7)}}{10^{0.18(H-15.7)} +
    10^{-0.18(H-15.7)}},
\label{eq:lsstmba}
\end{equation}
normalizing this curve to the observed $H<17$ counts for each region.
